---
title: "What strategies do ETH students use to minimise food waste during weekday lunches?"
subtitle: "Capstone Project"
author: "Fionn Smith"
affiliation: "ETH Zurich"
license: CC-BY-4.0
format: html
date: today
bibliography: "http://127.0.0.1:23119/better-bibtex/export/collection?/2/TheDeathrae.bibtex"
toc: true
execute: 
  echo: true
  warning: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(gt)

source(here::here("R/01-data_download.R"))
source(here::here("R/02-data_cleaning.R"))

```

# Introduction

Food waste... This report... @lealfilho2024food @pinto2018simple @whitehair2013written

# Methods

Data was collected using a self-administered online questionnaire created in Google Forms. Participants were surveyed by: direct in-person approach with a QR code on campus, posters with QR codes displayed on central university notice boards and digital distribution of links in smaller student groups. In all cases, the purpose of the project was explained beforehand and responses were submitted voluntarily without supervision.

# Results

```{r}

```

# Conclusions

# References
