---
title: "What strategies do ETH Zurich students use to minimise food waste during weekday lunches?"
subtitle: "Capstone Project"
author:
  name: "Fionn Smith"
  orcid: 0009-0005-0172-5024
  email: fsmith@student.ethz.ch
  affiliation: 
    name: ETH Zurich
    url: https://ethz.ch/de.html
license: "CC BY 4.0"
format: html
date: today
bibliography: "http://127.0.0.1:23119/better-bibtex/export/collection?/2/TheDeathrae.bibtex"
toc: true
execute: 
  echo: true
  warning: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(gt)
library(ggthemes)

source(here::here("R/01-data_download.R"))
source(here::here("R/02-data_cleaning.R"))

```

# Introduction

In 2019, global food production accounted for approximately 31% of worldwide greenhouse gas emissions @tubiellopre2022. At the same time, preventable food waste is highly prevalent in Europe @pinto2018simple. In Switzerland, households are responsible for 38% and gastronomy for 14% of eco-points (EPs) caused by preventable food waste @beretta2019lebensmittelverluste. Reducing food waste has the potential to decrease greenhouse gas emissions. To gain insight into consumer behaviour and strategies for food waste reduction, this report examines a survey conducted as part of the Research Beyond the Lab course in FS25 to analyse what strategies ETH Zurich university students use to minimise food waste during weekday lunches.

# Methods

Data was collected anonymously using a self-administered online questionnaire created in Google Forms. A total of 31 participants were surveyed by: (1) direct in-person approach, where a QR code for participation was presented to passing students outside on campus on a weekday during the semester over a span of 20 minutes; (2) posters with QR codes displayed on central university notice boards in buildings on both ETH Zurich campuses for 12 days; and (3) digital distribution of a link through a student-specific group chat. \
In all cases, the purpose of the project was explained beforehand, and responses were submitted voluntarily without supervision. The survey was open for responses for 12 days, excluding a prior testing period.

# Results

```{r}
survey_data <- read_csv(here::here("data/processed/university-lunch-food-waste_processed.csv"))
```

```{r}
consciousness_mean <- round(mean(survey_data$consciousness), 2)
```

The mean self reported consciousness of food waste among students is `r consciousness_mean` on a scale of 1 (not conscious at all) to 5 (very conscious). @fig-consciousness shows the distribution by study cycle. Master's students reported highest levels of consciousness, while Bachelor's and PhD students see a peak at 3/5. However, due to the low sample size, especially for PhD students, no clear conclusions can be drawn, leaving the correlation of self reported consciousness to study cycle requiring further invest

```{r}
#| label: fig-consciousness
#| fig-cap: "Consciousness of food waste by study cycle"

consciousness_data <- survey_data |> 
  select(consciousness, student_type) |> 
  mutate(student_type = case_when(
    student_type == "bachelor" ~ "Bachelor's",
    student_type == "master" ~ "Master's",
    student_type == "phd" ~ "PhD"
  ))

ggplot(data = consciousness_data, 
       aes(x = consciousness, 
           fill = student_type)) +
  geom_bar(col = "grey30") +
  theme_fivethirtyeight() +
  scale_color_colorblind() +
  labs(x = "Consciousness Level",
       y = "Count",
       fill = "Study Cycle")

write_csv(consciousness_data, here::here("data/final/plt-consciousness.csv"))
```

Among students sustainability was found overwhelmingly to be a motivation to care about food waste @tbl-motivation. This suggests that the environmental impact associated with food production is well known and considered an important issue.

```{r}
#| label: tbl-motivation
#| tbl-cap: "Motivation to care about food waste proportion"

motivation_data <- survey_data |> 
  select(motivation_sustainability, motivation_monetary, motivation_humanitarian, motivation_cultural) |> 
  summarise(across(everything(), ~mean(. == TRUE))) |> 
  pivot_longer(cols = everything(), names_to = "Motivation", values_to = "Yes") |>
  mutate(No = 1 - Yes) |> 
  mutate(Motivation = case_when(
    Motivation == "motivation_sustainability" ~ "Sustainability",
    Motivation == "motivation_monetary" ~ "Monetary",
    Motivation == "motivation_humanitarian" ~ "Humanitarian",
    Motivation == "motivation_cultural" ~ "Cultural"
  ))

motivation_data |> 
  knitr::kable(digits = 2)

write_csv(motivation_data, here::here("data/final/tbl-motivation.csv"))
```

The most common reported lunch source is bringing lunch from home to university, followed by eating in a cafeteria with buying externally clear lowest @tbl-lunch_source. When considering this in connection with @tbl-lunch_waste where the only sources reporting higher than "near none" on food waste are university cafeterias, university cafeteria food waste stands out as a potential point of improvement.

```{r}
#| label: tbl-lunch_source
#| tbl-cap: "Days of lunch source per week (0-5)"

lunch_source_data <- survey_data |> 
  filter(!is.na(days_uni)) |> 
  select(days_cafeteria, days_own, days_external) |> 
  summarise(across(everything(), ~mean(.))) |> 
  rename(
    "Cafeteria" = "days_cafeteria",
    "Brought from home" = "days_own",
    "External" = "days_external"
  )

lunch_source_data |> 
  knitr::kable(digits = 2)

write_csv(lunch_source_data, here::here("data/final/tbl-lunch_source_data.csv"))
```

```{r}
#| label: tbl-lunch_waste
#| tbl-cap: "Waste by lunch source (1-5)"

lunch_waste_data <- survey_data |> 
  filter(!is.na(days_uni)) |> 
  select(waste_cafeteria, waste_own, waste_external) |> 
  summarise(across(everything(), ~mean(.))) |> 
  rename(
    "Cafeteria" = "waste_cafeteria",
    "Brought from home" = "waste_own",
    "External" = "waste_external"
  )

lunch_waste_data |> 
  knitr::kable(digits = 2)

write_csv(lunch_waste_data, here::here("data/final/tbl-lunch_waste_data.csv"))
```

# Conclusions

# References
