---
title: "What strategies do ETH students use to minimise food waste during weekday lunches?"
subtitle: "Capstone Project"
author:
  name: "Fionn Smith"
  orcid: 0009-0005-0172-5024
  email: fsmith@student.ethz.ch
  affiliation: 
    name: ETH Zurich
    url: https://ethz.ch/de.html
license: "CC BY 4.0"
format: html
date: today
bibliography: "http://127.0.0.1:23119/better-bibtex/export/collection?/2/TheDeathrae.bibtex"
toc: true
execute: 
  echo: true
  warning: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(gt)

source(here::here("R/01-data_download.R"))
source(here::here("R/02-data_cleaning.R"))

```

# Introduction

In 2019 global food production accounted for approximately 31% of worldwide greenhouse gas emissions @tubiellopre2022. At the same time, preventable food waste is highly prevalent in Europe @pinto2018simple. In Switzerland households are responsible for 38% and gastronomy for 14% of eco-points (EPs) caused by preventable food waste @beretta2019lebensmittelverluste. Reducing food waste has the potential to decrease greenhouse gas emissions. To gain insight into consumer behaviour and strategies for food waste reduction, this report examines a survey conducted as part of the Research Beyond the Lab course in FS25 to analyse what strategies ETH Zurich university students use to minimise food waste during weekday lunches.

# Methods

Data was collected anonymously using a self-administered online questionnaire created in Google Forms. A total of 31 participants were surveyed by: (1) direct in-person approach where a QR code for participation was presented to passing students outside on campus on a weekday during the semester over a span of 20 minutes, (2) posters with QR codes displayed on central university notice boards in buildings on both ETH Zurich campuses for 12 days and (3) digital distribution of a link through a student-specific group chat. \
In all cases, the purpose of the project was explained beforehand and responses were submitted voluntarily without supervision. The survey was open for responses for 12 days, excluding prior testing period.

# Results

```{r}

```

# Conclusions

# References
