---
title: "What strategies do ETH Zurich students use to minimise food waste during weekday lunches?"
subtitle: "Capstone Project"
author:
  name: "Fionn Smith"
  orcid: 0009-0005-0172-5024
  email: fsmith@student.ethz.ch
license: "CC BY 4.0"
format: html
date: today
bibliography: "http://127.0.0.1:23119/better-bibtex/export/collection?/2/TheDeathrae.bibtex"
toc: true
execute: 
  echo: true
  warning: false
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(gt)
library(ggthemes)

source(here::here("R/01-data_download.R"))
source(here::here("R/02-data_cleaning.R"))

```

# Introduction

In 2019, global food production accounted for approximately 31% of worldwide greenhouse gas emissions @tubiellopre2022. At the same time, preventable food waste is highly prevalent in Europe @pinto2018simple. In Switzerland, households are responsible for 38% and gastronomy for 14% of eco-points (EPs) caused by preventable food waste @beretta2019lebensmittelverluste. Reducing food waste has the potential to decrease greenhouse gas emissions. To gain insight into consumer behaviour and strategies for food waste reduction, this report examines a survey conducted as part of the Research Beyond the Lab course in FS25 to analyse what strategies ETH Zurich university students use to minimise food waste during weekday lunches.

# Methods

Data was collected anonymously using a self-administered online questionnaire created in Google Forms. A total of 31 participants were surveyed by: (1) direct in-person approach, where a QR code for participation was presented to passing students outside on campus on a weekday during the semester over a span of 20 minutes; (2) posters with QR codes displayed on central university notice boards in buildings on both ETH Zurich campuses for 12 days; and (3) digital distribution of a link through a student-specific group chat.\
In all cases, the purpose of the project was explained beforehand, and responses were submitted voluntarily without supervision. The survey was open for responses for 12 days, excluding a prior testing period.

# Results

```{r}
survey_data <- read_csv(here::here("data/processed/university-lunch-food-waste_processed.csv"))
```

```{r}
consciousness_mean <- round(mean(survey_data$consciousness), 2)
```

The mean self-reported consciousness of food waste among students is `r consciousness_mean` on a scale of 1 (not conscious at all) to 5 (very conscious). @fig-consciousness shows the distribution by study cycle. Master's students reported highest levels of consciousness, while Bachelor's and PhD students see a peak at 3/5. However, due to the low sample size, especially for PhD students, no clear conclusions can be drawn, suggesting the need for further investigation into the correlation of self-reported consciousness of food waste to study cycle.

```{r}
#| label: fig-consciousness
#| fig-cap: "Self-reported consciousness of food waste by study cycle"

consciousness_data <- survey_data |> 
  select(consciousness, student_type) |> 
  mutate(student_type = case_when(
    student_type == "bachelor" ~ "Bachelor's",
    student_type == "master" ~ "Master's",
    student_type == "phd" ~ "PhD"
  ))

ggplot(data = consciousness_data, 
       aes(x = consciousness, 
           fill = student_type)) +
  geom_bar(col = "grey30") +
  theme_igray() +
  scale_color_colorblind() +
  labs(x = "Consciousness Level",
       y = "Count",
       fill = "Study Cycle")

write_csv(consciousness_data, here::here("data/final/fig-consciousness.csv"))
```

Sustainability was reported as a motivation to care about food waste by a large majority of students, more frequently than other reasons (@tbl-motivation). This suggests that the environmental impact associated with food production is widely recognised and considered an important issue.

```{r}
#| label: tbl-motivation
#| tbl-cap: "Motivation to care about food waste proportion"

motivation_data <- survey_data |> 
  select(motivation_sustainability, motivation_monetary, motivation_humanitarian, motivation_cultural) |> 
  summarise(across(everything(), ~mean(. == TRUE))) |> 
  pivot_longer(cols = everything(), names_to = "Motivation", values_to = "Yes") |>
  mutate(No = 1 - Yes) |> 
  mutate(Motivation = case_when(
    Motivation == "motivation_sustainability" ~ "Sustainability",
    Motivation == "motivation_monetary" ~ "Monetary",
    Motivation == "motivation_humanitarian" ~ "Humanitarian",
    Motivation == "motivation_cultural" ~ "Cultural"
  ))

motivation_data |> 
  knitr::kable(digits = 2)

write_csv(motivation_data, here::here("data/final/tbl-motivation.csv"))
```

In terms of lunch sources, bringing lunch from home to university was found to be most prevalent, followed by eating in a cafeteria. Buying lunch externally was clearly the least common choice (@tbl-lunch_source). When considering food waste the only sources reporting higher than "near none" on food waste are university cafeterias (@tbl-lunch_waste). This highlights university cafeteria food waste as a potential point of improvement.

```{r}
#| label: tbl-lunch_source
#| tbl-cap: "Days of lunch source per week (0-5)"

lunch_source_data <- survey_data |> 
  filter(!is.na(days_uni)) |> 
  select(days_cafeteria, days_own, days_external) |> 
  summarise(across(everything(), ~mean(.))) |> 
  rename(
    "Cafeteria" = "days_cafeteria",
    "Brought from home" = "days_own",
    "External" = "days_external"
  )

lunch_source_data |> 
  knitr::kable(digits = 2)

write_csv(lunch_source_data, here::here("data/final/tbl-lunch_source_data.csv"))
```

```{r}
#| label: tbl-lunch_waste
#| tbl-cap: "Waste by lunch source (1-5)"

lunch_waste_data <- survey_data |> 
  filter(!is.na(days_uni)) |> 
  select(waste_cafeteria, waste_own, waste_external) |> 
  summarise(across(everything(), ~mean(.))) |> 
  rename(
    "Cafeteria" = "waste_cafeteria",
    "Brought from home" = "waste_own",
    "External" = "waste_external"
  )

lunch_waste_data |> 
  knitr::kable(digits = 2)

write_csv(lunch_waste_data, here::here("data/final/tbl-lunch_waste_data.csv"))
```

To further investigate the reasons behind higher food waste in university cafeterias compared to other sources of lunch, students were presented with different scenarios and asked to choose their most likely approach to leftover food. Throwing food away was the least common choice, both in the case of eating lunch in a university cafeteria alone and eating with a friend. However, when eating alone, throwing leftovers away became more prevalent (@fig-scenarios). This suggests...

```{r}
#| label: fig-scenarios
#| fig-cap: "Leftover food scenarios"

scenario_data <- survey_data |> 
  select(scenario_lunch_alone, scenario_lunch_friend) |> 
  pivot_longer(cols = everything(), names_to = "Scenario", values_to = "Action") |>   mutate(
    Scenario = case_when(
      Scenario == "scenario_lunch_alone" ~ "Eating lunch alone",
      Scenario == "scenario_lunch_friend" ~ "Eating Lunch with a friend"
    ),
    Action = case_when(
      Action == "pass_on" ~ "Pass on",
      Action == "take_home" ~ "Take home",
      Action == "throw_away" ~ "Throw away"
  ))

ggplot(data = scenario_data, 
       aes(x = Action)) +
  facet_grid(~Scenario, scales = "free") +
  geom_bar(fill = "#377EB8", col = "grey30") +
  theme_igray() +
  scale_color_colorblind() +
  labs(x = "First choice",
       y = "Count")

write_csv(scenario_data, here::here("data/final/fig-scenarios.csv"))
```

```{r}
#| label: fig-encouragement
#| fig-cap: "Methods to encourage food waste reduction"

encouragement_data <- survey_data |> 
  select(encouragement) |> 
  rename("Encouragement" = "encouragement") |> 
  mutate(
    Encouragement = case_when(
      Encouragement == "tupperware" ~ "Own tupperware",
      Encouragement == "provided_single_use" ~ "Single-use packaging",
      Encouragement == "provided_tupperware" ~ "Tupperware provided",
      Encouragement == "none" ~ "None"
  ))

ggplot(data = encouragement_data, 
       aes(x = Encouragement)) +
  geom_bar(fill = "#377EB8", col = "grey30") +
  theme_igray() +
  scale_color_colorblind() +
  labs(x = "Factor",
       y = "Count")

write_csv(encouragement_data, here::here("data/final/fig-encouragement.csv"))
```

```{r}
#| label: fig-portions
#| fig-cap: "Portion choice in cafeterias"

portions_data <- survey_data |> 
  select(portion_small, portion_large, portion_second) |> 
  pivot_longer(cols = everything(), names_to = "Portion", values_to = "Choice") |> 
  mutate(Portion = case_when(
      Portion == "portion_large" ~ "Large portion",
      Portion == "portion_small" ~ "Small portion",
      Portion == "portion_second" ~ "Second portion",
  ))

ggplot(data = portions_data, 
       aes(x = Choice)) +
  geom_bar(fill = "#377EB8", col = "grey30") +
  facet_grid(~Portion, scales = "free") +
  theme_igray() +
  scale_color_colorblind() +
  labs(x = "Choice",
       y = "Count")

write_csv(encouragement_data, here::here("data/final/fig-portions.csv"))
```

```{r}
#| label: tbl-avoidance
#| tbl-cap: "Avoidance of dislikes in cafeteria food"

avoidance_data <- survey_data |> 
  select(food_avoidance) |> 
  count(food_avoidance) |> 
  rename("Avoidance" = "food_avoidance",
         "Count" = "n") |> 
  mutate(Avoidance = case_when(
    Avoidance == "yes" ~ "Yes",
    Avoidance == "sometimes" ~ "Sometimes",
    Avoidance == "no" ~ "No",
  ))

avoidance_data |> 
  knitr::kable(digits = 0)

write_csv(avoidance_data, here::here("data/final/tbl-avoidance.csv"))
```

# Conclusions

-   Students report high levels of consciousness of food waste, with the vast majority consideirng sustainability as a motivation

-   University cafeterias are a common sources of weekday lunch in university and lead to more food waste than other sources

-   

# References
